---
title: "Assignment 01"
author: "Jawad Adil - 3049429"
date: "3/29/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

# Getting the data from CSV file
```{r DiamondData}
#Import the dataset
DiamondData <- read.csv("C:/Users/jawad adil/Downloads/DiamondData.csv")

#Get a subset of data for testing
s <- sample(nrow(DiamondData), size=10000, replace = FALSE, prob = NULL)
s <- DiamondData[s, ]

#Getting original data again for final report
s<- DiamondData
```

## Task 1 - omit/replace NA/incorrect values
```{r Task1}

#omitting NA values
s<-na.omit(s)

#correcting spelling for cut attribute
s$cut[s$cut == "Very Geod"] <-"Very Good"


#limiting the carat to given range, if the carat is greater than given range,
#set it to maximum value
s$carat<-replace(s$carat,s$carat>5.01,5.01)
```
\newpage
## Task 2 - Get summary of variables and draw pie,bar char, histogram and scatter plot for appropriate variables
```{r Task2}
#Get the summary of variables
summary(s)

#separate the numeric data to get variance
numeric_data <- data.frame(s$price,s$carat,s$depth,s$table,s$x,s$y,s$z)
#get the variance of numeric variables
var(numeric_data)
```
\newpage
```{r Task2a}

#pie chart for cut variable
pie(table(factor(s$cut)),main = "Pie chart for CUT property",labels = paste(levels(factor(s$cut)),
    paste(", Total = ",table(s$cut))),col=rainbow(nrow(table(s$cut))))
```
\newpage
```{r Task2b}
#bar plot for the color variable
barplot <- barplot(table(factor(s$color)),col=rainbow(nrow(table(s$color))),
                   xlab = "Colors",ylab = "Frequency",ylim=c(0,12000))
#printing counts on the bars of bar chart
text(barplot,table(s$color)+500,paste(table(s$color)) ,cex=1)
grid(col = "black")
```
\newpage
```{r Task2c}
#histogram for depth variable
hist1 <- hist(s$depth,main = "Depth Histogram", ylab = "Frequency",xlab = "Depth",
              ,ylim=c(0,30000),col=rainbow(10))
#printing counts on the histogram
text(hist1$mids+0.8,hist1$counts+3000,labels = hist1$counts,adj=c(0.5, -0.5),srt=90)
grid(col = "black")
```
This histogram is showing normality. Count of each value is shown on the top of bars.
\newpage
```{r Task2d}
#scatter plot for carat vs depth variables
plot(s$carat,s$depth,xlab = "Carat",ylab = "Depth",pch=20,col="#3299a8")
abline(lm(s$depth~s$carat),col="#9c32a8",lwd="3")
grid(col = "black")
#finding correlation
cor(s$carat,s$depth)
```
This Graph is showing normality and the relationship among them is very weak. Correlation function also results the same.
\newpage

## Task 3:

### 3A: histogram of price variable

```{r Task3a}
#histogram of price variable
histPrice <- hist(s$price,main = "Price Histogram", ylab = "Frequency",xlab = "Price",
     col=rainbow(10),xlim=c(0,20000),ylim=c(0,15000))
#printing counts on the bars of histogram
text(histPrice$mids+500,histPrice$counts+1500,labels = histPrice$counts,adj=c(0.5, -0.5),srt=90)
#printing the summary on the histogram using text function
text(15000,12000,paste("Median:",summary(s$price)[3]),pos=3,col="Blue")
text(15000,11000,paste("Mean:",as.integer(summary(s$price)[4])),pos=3,col="Blue")
text(15000,10000,paste("Variance:",as.integer(var(s$price))),pos=3,col="Blue")
text(15000,9000,paste("Max:",summary(s$price)[6]),pos=3,col="Blue")
text(15000,8000,paste("Min:",summary(s$price)[1]),pos=3,col="Blue")
grid(col = "black")
```
\newpage

### 3B: Group Diamonds by some price ranges and summerize them separately
```{r Task3b}
#divide the price variable into 3 groups

#low are those values which are less than 1st Quartile
low <- subset(s,price<=summary(s$price)[2])

#med are those values which are less than 3rd Quartile
med <- subset(s,price<=summary(s$price)[5] & price>summary(s$price)[2])

#all other greater values are high values
high <- subset(s,price > summary(s$price)[5])

#printing the count of low values
nrow(low)
#summary of low variable values
summary(low)
```
\newpage
```{r Task3bb}
#printing the count of med values
nrow(med)
#summary of med variable values
summary(med)
```
\newpage
```{r Task3bbb}
#printing the count of high values
nrow(high)
#summary of high variable values
summary(high)
```
\newpage

### 3C: Exploring different cut types using boxplot
```{r Task3c}
#Price vs Cut box plot
boxplot(s$price~s$cut,main="Price vs Cut distribution",xlab = "Cut",ylab = "Price",
        col=rainbow(nrow(table(s$cut))))
grid()
```
\newpage

### 3D: Finding relation of attributes with price
```{r Task3D}
#separate numeric data from actual dataset so we can directly use cor() function
numeric_data <- data.frame(s$price,s$carat,s$depth,s$table,s$x,s$y,s$z)

#finding the correlation between variables.
cor(numeric_data)
```
Price is correlated most with the X variable, Y is the 2nd high correlated variable with price and Z is the 3rd most correlated variable. Simply, the order of correlation is: 

### Most Strong: 
#### X variable 
### Mid Strong: 
#### Y variable 
### Least Strong: 
#### Z variable
#### All other variables have less strong relationship with price than these variables.

\newpage

## Task 4:

### 4A: compute the volume & plot price vs volume graph
```{r Task4a}
#separating x, y, z variables from dataset for better understanding
x <- s[,8]
y <- s[,9]
z <- s[,10]

#calculating volume variable
volume <- x*y*z

#binding old dataset and volume and storing them in new variable so old one
#remain unchanged
newData=cbind(s,volume)

#getting price from new dataset
price=newData[,7]

#ploting graph between price & volume
plot(volume,price,xlab = "Volume",ylab = "Price", main = "Price vs Volume distribution",
     pch=20,col="#3299a8")
abline(lm(price~volume),col="#9c32a8",lwd="3")
grid(col="black")

#finding correlation between price and volume
cor(price,volume)
```
The graph shows direct and very strong relationship between price and volume, strength is also proven by correlation function.
\newpage

### 4B: Finding correlation between carat and volume & plotting them with regression line

```{r Task4b}
#plotting the graph between carat and volume
plot(newData$carat,volume,xlab = "Carat",main = "Carat vs Volume distribution",
     pch=20,col="#3299a8")

#drawing a regression line on graph
abline(lm(newData$carat ~ volume),col="#9c32a8",lwd="3")
grid(col="black")

#finding correlation between volume and carat
cor(newData$carat, volume)

```
The graph shoes the strong correlation between volume and carat which is also proved by cor() function.
\newpage

### 4C: Exploring relationship of table and depth variables:

```{r Task4C}
#finding correlation between table and depth variables
cor(newData$depth,newData$table)

#plotting the relationship graph
plot(newData$depth,newData$table,xlab = "Depth",ylab = "Table",
     main = "Depth vs Table distribution",pch=20,col="#3299a8")

#drawing regression line
abline(lm(newData$depth~newData$table),col="#9c32a8",lwd="3")
grid(col = "black")
```
The graph shows negative relationship which is also proven by the cor() function. 
\newpage

### 4D: Exploring relationship of table and rest of the variables:
```{r Task4D}
#ploting relationship between price and table
plot(newData$price,newData$table,xlab = "Price",ylab = "Table",
     main = "Price vs Table distribution",pch=20,col="#3299a8")

#drawing regression line
abline(lm(newData$price~newData$table),col="#9c32a8",lwd="3")
grid(col = "black")

#finding correlation between price and table
cor(newData$table,newData$price)
```
The graph between price and table shows weak but positive relationship which is also proven by the cor() function. 
\newpage
```{r Task4Dd}
#plotting carat vs table relationship
plot(newData$carat,newData$table,xlab = "Carat",ylab = "Table"
     ,main = "Carat vs Table distribution",pch=20,col="#3299a8")

#plotting regression line
abline(lm(newData$table~newData$carat),col="#9c32a8",lwd="3")
grid()

#finding correlation between table and carat
cor(newData$table,newData$carat)

```
The Graph show weak but positive relationship between carat and table which is also proven by correlation function. Also, this graph is showing normality.

\newpage
```{r Task4Ddd}
#plotting carat vs table relationship
plot(newData$volume,newData$table,xlab = "Volume",ylab = "Table"
     ,main = "Volume vs Table distribution",pch=20,col="#3299a8")

#plotting regression line
abline(lm(newData$table~newData$volume),col="#9c32a8",lwd="3")
grid()

#finding correlation between table and carat
cor(newData$table,newData$volume)

```
The Graph between volume & table also shows a weak but positive relation which is also proven by correlation function. This graph is also showing normality as the line is going slighlty upward as we move right from left.
